---
title: "8/30/18 notes"
author: "Brennan Donnell"
output: pdf_document
---


# NOTES
## Intro

Replication is important when writing code. Make sure it's legible and easily understandable, and that your variables make sense. 
you also want to show what worked **and** what didn't. 

Side note, go back and redo the code in python/r and write a report about it. You could get it published too. Go to a conference with it as opposed to a journal. 

## R Markdown
What is it? 

it's markdown mixed with R code. Its output is similar to HTML, but way easier to write. 
Good thing about it is that you can run your R code **AND** your python code. 

# THIS IS HEADING 1
## this is a second heading. 
From the pop-up window, if you select **Open in browser** it opens in your browser. 

## lists and asterisks

* Putting *one* asterisk around text gives you *italicized* text. 
* Putting **two** asterisks around text gives you **bolded** text. 
    + To give you a list, give a space between the start of the list

[link](www.rstudio.com) is how you make a link. 

***
# Example: mexample.txt
## create the stuff in r markdown. 
Applicants must hold a baccalaureate degree (no specific field of study is required) from an accredited college or university with a **3.0 GPA** or higher. Applicants should have demonstrated quantitative aptitude in undergraduate coursework or similar work experience in analytic field. Prior coursework may include: ***statistics, chemistry, physics, mathematics*** or other quantitative courses.
Submit an application to the [UNH Graduate School](http://gradschool.unh.edu/apply.php) with the following:

1. Prior college transcripts
2. Two letters of recommendation
3. Resume

The document was last update on: `r Sys.time()`

***
# How about actually using R code in this? 
## simple. 
Here's how it is done. This is called an **R Chunk**
The following function (*paste*) concatenates two strings: "AB" and "CD". 
```{r}
fn <- "AB"
ln <- "CD"
print(paste(fn, ln))

x <- "Univ of"
y <- "New Hampshire"
paste(x, y)
```
Now let's say that this code didn't work, but we still want to keep it. How do we hide it? in the R chunk, put {r echo=FALSE}. 
```{r echo=FALSE}
fn <- "AB"
ln <- "CD"
print(paste(fn, ln))

x <- "Univ of"
y <- "New Hampshire"
paste(x, y)
```
But now, we're still getting the output. WE don't even want that. Now try {r echo=FALSE, eval=FALSE}. With the new eval=FALSE, the code in the chunk doesn't even get executed. 
```{r echo=FALSE, eval=FALSE}
fn <- "AB"
ln <- "CD"
print(paste(fn, ln))

x <- "Univ of"
y <- "New Hampshire"
paste(x, y)
```
To turn off warnings, such as from loading libraries, if you include warning=FALSE in the curly braces for the chunk, the warnings won't appear. 

Lastly, to showcase the code but *not* the results, use {r results='hide'} in the chunk. 

***
# question
## in R markdown...
use data(mtcars) and do some stuff
SCATTERPLOT of **MPG vs HP**
```{r warning=FALSE, results='hide', mtcars, fig.height=4}
library(ggplot2)
#plot(mtcars$mpg, mtcars$hp)
ggplot(mtcars, aes(hp, mpg)) + geom_point(aes())
```
Bar Pot of **gear**
```{r}
ggplot(mtcars, aes(gear)) + geom_bar()
```
***

# BENCHMARKING

## WHY DO WE NEED IT? 

Is your code **too slow**? Does your system **Crash** whenever you run R?
Well, *try keeping your libraries up-to-date*. 
system.time(command) times it. 
```{r warning=FALSE}
library(readr)
```
which runs faster, read.csv() or read_csv()?
```{r warning=FALSE}
setwd("/Users/brennandonnell/Desktop/grad_school/R")
system.time(
  read.csv("NH_data_072418.csv")
  )

system.time(
  read_csv("NH_data_072418.csv")
)
```

check out the library(microbenchmark). this'll do it a lot better. 

```{r warning=FALSE, eval=FALSE}
library(microbenchmark)
microbenchmark(read.csv("NH_data_072418.csv"),
               read_csv("NH_data_072418.csv"),
               times = 3)
```

an important thing to know for parallel processing is *how many cores does your computer have*? 
Luckily, thereis an easy way to figure it out, thanks to the benchmarkme library
```{r}
library(benchmarkme)
get_ram()
get_cpu()
```
# Code

* Which part of the code is taking the longest time to execute? 
* Can you change the way you are executing the code? 
    + Which is taking up the most time? 

this is where the library profvis comes in. 

* It allows you to benchmark *each line* of our code for...
    + d
    + d

how do you use it? 
profvis({
your code here
})

```{r warning=FALSE, eval=FALSE}
library(profvis)
profvis({
  library(readr)
  x <- read_csv("Durham_Weather.csv")
  x$below_zero_vector <- ifelse(x$TMIN < 0, 1, 0)
  for (i in 1:length(x$TMIN)){
    if (x$TMIN[1] < 0){
      x$below_zero_loop[i] <- 1
    }else{
      x$below_zero_loop[i] <- 0
    }
    }
})
```
 
*** 
write code to make a scatterplot of tmin vs tmax using ggplot and plot. According to profvis, which is better? 

```{r warning=FALSE, eval=FALSE}
profvis({
   x <- read_csv("Durham_Weather.csv")
   durham_weather <- filter(x, !is.na(TMAX) | !is.na(TMIN))
   plot(durham_weather$TMIN, durham_weather$TMAX)
   ggplot(durham_weather, aes(TMIN, TMAX)) + geom_point()
})
```
All odd occurances aside, it
what does *version* do? `r version`
it prints out the version & other stuff. 

***

# Question: 

* Does every year have 365 days? (no, leap year)
* Does every day have 24 hours? (no, dst)
* Does every minute have 60 seconds? (no, leap second. )

SO... what is the right day or time? what date and time is universal? what is the global standard time? [ISO 8601](https://en.wikipedia.org/wiki/ISO_8601) is the golden standard. 

* yyyy-mm-dd
    + Example: 24 July 2018 in ISO 8601 is 2018-07-24, no need of separators (or - if needed). 
* HH:MM:SS
    + Example: 9.15 PM in ISO 8601 standard is 21:15:00
```{r}
x <- Sys.Date()
```

*** 

# fun

Use the appropriate lubridate function to parse each of the following dates: 

* d1 <- "August 29, 2018"
* d2 <- "2018-Aug-29"
* d3 <- "29-Aug-2018"
* d4 <- c("August 29 (2018)","August 30 (2018)")
* d5 <- "08/29/18"

(if you don't remember lubridate, google the following: DMY, MDY, YMD)

```{r}
library(lubridate)
d1 <- mdy("August 29, 2018")
d2 <- ymd("2018-Aug-29")
d3 <- dmy("29-Aug-2018")
d4 <- mdy(c("August 29 (2018)", "August 30 (2018)"))
d5 <- mdy("08/29/18")
```

Now try reading in durham weather again and telling the class of the variable

```{r}
y <- read.csv("Durham_Weather.csv")
typeof(y$Temp_Date)

x <- read_csv("Durham_Weather.csv")
typeof(x$Temp_Date)
typeof(x$Raw_Date)
```
SO, read.csv gives us an integer, and read_csv gives us a double. better. 
now, extract and create new columns for

* day
* month (label)
* year
* weekday (label)
* yearday 

(hints: In lubridate, day, month, year, wday, yday)
```{r warning=FALSE}
x1 <- read_csv("Durham_Weather.csv")
x1$Day     <- day(x$Temp_Date)
x1$Month   <- month(x$Temp_Date, label = TRUE)
x1$Year    <- year(x$Temp_Date)
x1$Weekday <- wday(x$Temp_Date, label = TRUE)
x1$Yearday <- yday(x$Temp_Date)
```

Now try doing that with the rawdate. 
```{r warning=FALSE}
x2 <- read_csv("Durham_Weather.csv")
x2$Day     <- day(x$Raw_Date)
x2$Month   <- month(x$Raw_Date, label = TRUE)
x2$Year    <- year(x$Raw_Date)
x2$Weekday <- wday(x$Raw_Date, label = TRUE)
x2$Yearday <- yday(x$Raw_Date)
```
INterestingly enough, the date in Raw_Date switches midway through, resulting in a lot of missing data. How do we fix this? 

d2$raw_date_fixed <- parse_date_time... just check the notes. 

***

# Work on the following before coming back: 

## checking for holidays? 

* for the Durham_Weather dataset, create a column "isholiday" which equals 1 if NYSE is closed, or 0 if NYSE is open for the years 2014 & 2015
* SOme of the functions of interest to solve this problem include...
    + holidayNYSE from library(timeDate)
    + %in%

holidayNYSE takes in a year and gives back the timedates of the holidays
```{r warning=FALSE, message=FALSE}
library(timeDate)
hw           <- read_csv("Durham_Weather.csv")
hw$Day       <- day(hw$Temp_Date)
hw$Month_lab <- month(hw$Temp_Date, label = TRUE)
hw$Month     <- month(hw$Temp_Date)
hw$Year      <- year(hw$Temp_Date)
hw$Weekday   <- wday(hw$Temp_Date, label = TRUE)
hw$Yearday   <- yday(hw$Temp_Date)
hw$YMD       <- ymd(paste(hw$Year, hw$Month, hw$Day, sep = "-"))
hw$isweekend <- ifelse(hw$Weekday %in% c("Sat", "Sun"), 1, 0)
hw$isholiday <- ifelse(hw$YMD %in% as.Date(holidayNYSE(2014:2015)), 1,
                       ifelse(hw$isweekend == 1, 1, 0))
```
