---
title: "9/6/2018 notes"
author: "Brennan Donnell"
date: "9/6/2018"
output: html_document
runtime: shiny
---
```{r libraries}
library(dplyr)
library(ggplot2)
```
# Using python in RMD
The library reticulate enables to use python with r, you also have to direct to where it is using the *use_python* method. 
```{r python_enabler}
library(reticulate)
use_python("/Users/brennandonnell/anaconda/bin/python")
```
Printing the first ten natural numbers numbers using r
```{r r_code}
x <- 1:10
print(x)
```
Printing the first ten natural numbers using python
```{python python_code}
import numpy as np
y = np.arange(1,11)
print(y)
```
# Shiny: 
## What is it? 
This is what you use if you can't afford tableau and/or you are on a budget. 

* R package for *interactive web apps*

## 3 major road blocks
1. **User INterface (UI) aka look and feel**
    + Defines the components of the user interface
        + Page titles
        + Input options
        + Outputs
    + Defines what the user will see and interact with
2. **Server, aka brain**
    + Contains the information to build the app
    + Defines what happens in R
3. **Global/shinyapp aka enabler**

```{r shiny_example_1}
library(shiny)
# User Interface
ui_demo <- fluidPage(
 # "UNH Analytics & Data Science"
  #
  tags$em(h2("UNH Analytics & Data Science"))
  )
# Server
server_demo <- function(input, output){
}

# Enabler, creates the app. 
shinyApp(ui=ui_demo, server=server_demo)
```
You can also use tags for this, see the comments. 
```{r html_tags}
names(tags)
```
## Making user input

```{r shiny_example_2}
ui2 <- fluidPage(
  sidebarLayout(
    sidebarPanel(textInput(inputId = "n1", label = "input name")), 
    mainPanel(textOutput(outputId = "displaytext"))
    )
  )

server2 <- function(input, output) {
  output$displaytext <- renderText(
    print(paste("Hello", input$n1))
  )
}
shinyApp(ui=ui2, server = server2)
```
## another example
make it do it for two strings together
```{r shiny_example_3}
ui3 <- fluidPage(
  sidebarLayout(
    sidebarPanel(textInput(inputId = "n1", label = "input first"),
                 textInput(inputId = "n2", label = "input second")), 
    mainPanel(textOutput(outputId = "displaytext"))
    )
  )

server3 <- function(input, output) {
  output$displaytext <- renderText(
    print(paste(input$n1, input$n2))
  )
}
shinyApp(ui=ui3, server = server3)
```
## make a calculator
doesn't work yet
```{r shiny_example_4}
ui4 <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      numericInput(inputId = "n1", label = "First number", 0),
      numericInput(inputId = "n2", label = "Second number", 0)), 
    mainPanel(textOutput(outputId = "calculate"))
    )
  )

server4 <- function(input, output) {
  output$calculate <- renderText(
    sum(input$n1, input$n2)
  )
}
shinyApp(ui=ui4, server = server4)
```
# Moving on to the csv

```{r reading_csv, warning=FALSE, error=FALSE, message=FALSE}
library(readr)
fastfood_data <- read_csv("/Users/brennandonnell/Desktop/grad_school/R/fast_food (1).csv")
```
Doing stuff in shiny!

```{r radiobuttons}
ui_food <- fluidPage(
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    # Sidebar panel for inputs ----
    sidebarPanel(
      radioButtons("chain_ui", "Select Chain", choices = unique(fastfood_data$Chain), 
                         selected = unique(fastfood_data$Chain)[1],inline = FALSE)
      
    
      ),
    # Main panel for displaying outputs ----
    mainPanel(
      # Output: Text ----
      h2("Scatter Plot"),
      plotOutput(outputId = "scatter")
      )
  )
)
server_food <- function(input,output){
  filtered_data <- reactive(filter(fastfood_data, Chain == input$chain_ui))
  output$scatter <- renderPlot(
    {
      ggplot(filtered_data(), aes(x=Calories,y=Saturated_Fat))+ geom_point()+
                    theme_classic()+scale_x_continuous(limits = c(0, 1500)) +
                    scale_y_continuous(limits = c(0,35))
    }
  )
}

shinyApp(ui = ui_food, server = server_food)
```

#What if we want to make a server of it? 
Copyt he following into a new folder.
```{r za_warudo}
library(shiny)
library(dplyr)
library(ggplot2)
library(readr)
fastfood_data <- read_csv("fast_food.csv")
```
in the previous cell, make separate .r scripts for the server, ui, and the above will be the global

---
#shinyapps.io stuff
Go to [shinyapps.io](https://www.shinyapps.io) and make an account. It's pretty nifty and will let you share your apps. You can also link it up to your GitHub account, so you should also do that. 

Go to Account, and then token. Don't share your token. 
Now, in RStudio, under *Tools*, select *Global Options*, and then pick *Publishing*. You'll be able to *Connect* your shinyapps.io account to this, by putting the super secret code into there. 

## This is where the fun begins. 
Now, open one of your global files. a little button that says *Run app* will appear From there it will do the thing. **But wait, there's more!** If you click on the *publish* icon, you can publish it to your shinyapps.io account, and then run it on that website, and other people can also run it. 

**Note:** whenever you publish stuff, you don't need the enabler. 

---
# For next week...

* Create multiple tabs: one with a barplot, another with a scatterplot
* get your internet to work
* make a shinyapps.io account
    + Hint: use tabsetPanel() and tabPanel() as necessary within the mainPanel. 
    
ON to tableauprep